---
title: "HW_9_Q4"
author: "Colin Frishberg"
output: pdf 
---

```{r}
library(tidyverse)
library(sandwich)
library(lmtest)
library(stargazer)
library(fec16)
library(patchwork)

```
#Q4.1 Exploring - Make Histograms

```{r}
#Observe Data
head(campaigns)
head(results_house) 


#Hist 1 
p1 <- ggplot(results_house, aes(x = general_percent)) +
  geom_histogram( aes(y=after_stat(density)),
                        col = "black",
                        fill = "darkblue") +
stat_function(fun = dnorm, args = list(mean = mean(results_house$general_percent, na.rm = TRUE), sd = sd(results_house$general_percent, na.rm = TRUE)), 
              col = "red")

#Hist 2 
p2 <- ggplot(campaigns, aes(x = ttl_disb)) +
  geom_histogram( aes(y = after_stat(density)),
                        col = "black",
                        fill = "darkblue") +
stat_function(fun = dnorm, args = list(mean = mean(campaigns$ttl_disb, na.rm = TRUE), sd = sd(campaigns$ttl_disb, na.rm = TRUE)),
              col = "red")

#Patchwork to create a side by sides comparison of the two histograms
p1 + p2
```

#Q4.2/3 Exploring - Build DF parts 1 and 2

```{r, include=FALSE}
#Join on cand_id using DPLYR
df1 <- inner_join(results_house, campaigns, by = "cand_id")

#Count rows and columns
nrow(df1)
ncol(df1)
```
#Q4.4 Exlporing - Indicator Variables and Scatter Plot

```{r}
#Find the Right Column
head(df1)

#Identify the Different Values in the "Party" Column 
unique(df1$party)

#general_votes = # of votes in the general election in this case for the house of representatives
#candidate_party = Democratic, Republican, Other
# ttl_disb = total_disbursements 


df1 <- df1 %>% 
  mutate(
    candidate_party = case_when(
      party == "REP" ~ "Republican",
      party =="DEM" ~ "Democrat",
      TRUE ~ "Other Party"
    ) 
  )



#Create A Scatter plot 
ggplot(df1, aes(x = ttl_disb, y = general_votes, colour = candidate_party)) + 
  geom_point(alpha = .3, na.rm = TRUE) +
  scale_colour_manual(values = c("Democrat" = "blue",
                                 "Republican" = "red",
                                 "Other Party" = "darkgreen"),
                      name = "Candidate Party") +
  labs(title = "2016 House Campaign Disbursments and Candidate Votes",
       x = "Disbursements",
       y = "Votes")
  
   

```

#Q4.5 Regression - Evaluate large sample assumptions

#Cleaning
```{r}
# Create a data frame with the 3 variables we are looking to fit the model and clear the data

reg_df <- df1 %>% select(general_votes,candidate_party,ttl_disb)

# Replace "Unopposed" with 0
reg_df$general_votes[reg_df$general_votes == "Unopposed"] <- 0

# Convert general_votes to numeric
reg_df$general_votes <- as.numeric(reg_df$general_votes)

# Clean the data frame
reg_df <- reg_df %>%
  filter(!is.na(general_votes), !is.na(candidate_party), !is.na(ttl_disb))

# Convert candidate_party to factor
reg_df$candidate_party <- as.factor(reg_df$candidate_party)

# Trim whitespace
reg_df$candidate_party <- trimws(reg_df$candidate_party)

str(reg_df)
levels(reg_df$candidate_party)



# Perform linear regression with multiple variables
model <- lm(general_votes ~ candidate_party + ttl_disb, data = reg_df)


#Model Summary
summary(model) 

## Note - it appears R has chose to use Democrats as the reference variable through the one-hot enconding method. 
```

#Assumptions tests

```{r}
#Coefficients
coef(model)

#Fitted values
hist(fitted(model))

summary(fitted(model))

#Residuals
resid(model)
hist(resid(model))
```

#Diagnostic plots
```{r}
plot(model, which = 1) #Residuals against Fitted Values
```

```{r}
plot(model, which = 2) # Normal Q-Q Plot
```

```{r}
#Breusch-Pagan test for heteroscedasticity 
bptest(model)
# p-value of .0181 leads me to reject the null hypothesis of heteroscedasticity 
```

```{r}
#Durbin Watson test for autocorrelation
dwtest(model)
```

```{r}
#To run t-tests for all coefficients using robust standard errors
coeftest(model, vcov = vcovHC)

```


##Examine the variable relationships to help with assumptions 
```{r}

#Categorical (Party) vs. outcome

plot_a <- ggplot(reg_df, aes(x = candidate_party, y = general_votes, fill = candidate_party)) +
  geom_boxplot() +
  scale_fill_manual(values = c("Democrat" = "blue",
                                 "Republican" = "red",
                                 "Other Party" = "darkgreen"),
                      name = "Candidate Party") +
  labs(title = "Party and Votes",
       x = "Candidate Party",
       y = "Votes")


#Continuous (ttl_disb) vs. outcome

plot_b <- ggplot(reg_df, aes(x = ttl_disb, y = general_votes)) +
  geom_point(alpha = .4) +
  scale_x_continuous() +
  scale_y_continuous() +
  labs(title = "Total Disbursement and Votes",
       x = "Total Disbursement",
       y = "General Votes")




(plot_a) | (plot_b)
```






```{r}
#Plot linear models against our sample to check if a linear model looks to exist 
ggplot(reg_df, aes(x = ttl_disb, y = general_votes, colour = candidate_party)) + 
  geom_point(alpha = .3) +
  geom_smooth(method = "lm", se = FALSE) +
  scale_colour_manual(values = c("Democrat" = "blue",
                                 "Republican" = "red",
                                 "Other Party" = "darkgreen"),
                      name = "Candidate Party") +
  labs(title = "2016 House Campaign Disbursments and Candidate Votes",
       x = "Disbursements",
       y = "Votes")
  
   
```

# Arguments about the large-sample assumption
1. IID 

For the first part of IID, we are looking to argue that the residuals are independent from each other so we can use a Durbin Watson test to test for autocorrelation. Our resluts `DW = 1.4533, p-value = 2.317e-16` suggest that we reject the null of no positive autocorrelation and given DW<2 we can infer that there is indeed positive autocorrelation in our model. To check for Identical Distribution, we observe the results of the QQ plot from `plot(model, which = 2)` suggest that the "identically distributed" assumption of IID is violated as the residuals are not normally distributed due to the presence of heavy tails and significant outliers. 


2. Unique BLP Exists 
We can use the Fitted Values vs. Residuals visual `plot(model, which = 1)` to check for linearity. Observing this plot, we see both a "Funnel" characteristic of the residuals, strong outliers and a distinct curve in our model. These visual elements along with the assumptions violated for IID incline me to suggest that a unique BLP does not exist for this particular set of variable relationships. 

#Q4.6 Stargazer 

```{r results = 'asis'}


stargazer(model, 
          type = "latex", 
          title = "Regression Results: General Votes", 
          dep.var.labels = "General Votes", 
          covariate.labels = c("Other Party", "Republican", "Total Disbursements"),
          notes = "Significance levels: * p < 0.05, ** p < 0.01, *** p < 0.001",
          align = TRUE,
          no.space = TRUE,
          digits = 3,
          out = "regression_table.tex" #save to .tex file
          )

```
#Q4.7 Regression - Money's Relationship with Votes


```{r}
model_1 <- lm(general_votes ~ ttl_disb, data = reg_df)
summary(model_1)


```
# Argument

Based on the data available, I would reject the null hypothesis that Total Disbursements has no relationship with Votes and conclude that there is a statistically significant relationship between Disbursements and Votes. Based on the R-squared value of .050, we see that the model explains around 50% of the variance in general_votes, which suggests that Total Disbursements had a meaningful impact on the number of votes a candidate received. 

#Q4.8 Regression - Party's Relationship with Votes

```{r}
model_2 <- lm(general_votes ~ candidate_party, data = reg_df)
summary(model_2)
```
# Argument 
Based on the data available I would reject the null hypothesis that Candidate Party has no relationship with Votes and conclude that there is a statistically significant relationship between a Candidate's Party and how many votes they receive. The R-squared value of .31 suggests that 31% of the variance in votes is explained by our model which does provide us a meaningful insight which suggests Total Disbursment is a stronger predictor of candidate votes than both candidate party, and a model containing them both. 

